**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-29 Schloss_PacBio16S_PeerJ_2016.Rmd
Extract the analysis of the error rates by each method

## Basic error correction analysis
get_raw_error <- function(region){
	region_data <- mock_error_good[mock_error_good$region == region,]
	return(100*sum(region_data$mismatches)/sum(region_data$total))
}

raw_rates <- sapply(regions, get_raw_error)

get_basic_kept <- function(region){
	region_data <- mock_error_good[mock_error_good$region == region,]
	100 * sum(region_data$good_coords & region_data$good_homop)/nrow(region_data)
}

basic_kept <- sapply(regions, get_basic_kept)

get_basic_error <- function(region){
	region_data <- mock_error_good[mock_error_good$region == region & mock_error_good$good_coords & mock_error_good$good_homop,]
	return(100*sum(region_data$mismatches)/sum(region_data$total))
}

basic_rates <- sapply(regions, get_basic_error)

mock_error_basic <- mock_error_good[mock_error_good$good_coords & mock_error_good$good_homop,]
basic_n <- aggregate(mock_error_basic$error, by=list(mock_error_basic$region), length)$x
names(basic_n) <- names(basic_rates)

basic_error_rate <- 100 * sum(mock_error_basic$mismatches) / sum(mock_error_basic$total)		#[1] 0.6448736

@@

```{r barcode_primer_analysis}
mock_error_basic$pbdiffs <- mock_error_basic$fbdiffs + mock_error_basic$rbdiffs + mock_error_basic$fpdiffs + mock_error_basic$rpdiffs

@@

# want to get the error rate and fraction of sequences for each region with a
# threshold of mismatches or fewer

pb_by_region_mm <- aggregate(mock_error_basic$error, by=list(mock_error_basic$pbdiffs, mock_error_basic$region), function(x){c(mean=mean(100*x), nseqs=length(x))})

aggregate_pb_data <- function(region){
	pb_region <- pb_by_region_mm[pb_by_region_mm$Group.2 == region,]

	product <- pb_region$x[,"mean"] * pb_region$x[,"nseqs"]
	pb_region_nseqs <- cumsum(pb_region$x[,"nseqs"])
	pb_region_error <- cumsum(product) / pb_region_nseqs
	pb_region_frac <- pb_region_nseqs / sum(pb_region$x[,"nseqs"])
	c(error=pb_region_error, fraction=pb_region_frac)
}

pb_data <- sapply(regions, aggregate_pb_data)

pb_error <- pb_data[grep("error", rownames(pb_data)),]
rownames(pb_error) <- 0:6

pb_fraction <- pb_data[grep("fraction", rownames(pb_data)),]
rownames(pb_fraction) <- 0:6

pb_error_reduction_0 <- 100 * (1 - pb_error["0",]/basic_rates)
pb_error_reduction_1 <- 100 * (1 - pb_error["1",]/basic_rates)

pb_nseqs_reduction_0 <- 100 * (1 - pb_fraction["0",])
pb_nseqs_reduction_1 <- 100 * (1 - pb_fraction["1",])
```

```{r coverage_analysis}
#This is for the Figure 2C plot...
coverage_overall_agg <- aggregate(100 * mock_error_basic$error, by=list(mock_error_basic$coverage), mean)
coverage_overall <- coverage_overall_agg$x
names(coverage_overall) <- coverage_overall_agg$Group.1
coverage_overall <- coverage_overall[as.character(3:40)]

# need to get the error rate for coverage values between 3 and 40 for each region

error_by_coverage <- function(region){
	coverage_region <- mock_error_basic[mock_error_basic$region == region,]
	coverage_error_agg <- aggregate(100 * coverage_region$error, by=list(coverage_region$coverage), mean)

	coverage_error <- rep(NA, max(mock_error_basic$coverage))
	coverage_error[coverage_error_agg$Group.1] <- coverage_error_agg[,"x"]
	coverage_error
}

coverage_error <- sapply(regions, error_by_coverage)
#seems to plateau right around 10

error_at_below_coverage_cutoff <- function(region, cutoff=10){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	deep <- region_data$coverage >= cutoff
	coverage_data <- region_data[deep,]

	error <- 100 * sum(coverage_data$mismatches)/sum(coverage_data$total)
	frac_kept <- nrow(coverage_data) / nrow(region_data)
	c(error=error, fraction=frac_kept)
}

coverage_data <- sapply(regions, error_at_below_coverage_cutoff)

coverage_error <- coverage_data["error",]
coverage_fraction <- coverage_data["fraction",]

coverage_error_reduction <- 100 * (1 - coverage_error/basic_rates)
coverage_nseqs_reduction <- 100 * (1 - coverage_fraction)
```

```{r predicted_error_analysis }
#This is for the Figure 2D plot...
pred_overall <- aggregate(100 * mock_error_basic$error, by=list(round(mock_error_basic$pred_error, 3)), mean)

error_at_above_pred_error_cutoff <- function(region, threshold=0.999){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	above_threshold <- region_data[region_data$pred_error >= threshold,]

	c(error=100*sum(above_threshold$mismatches)/sum(above_threshold$total), nseqs=nrow(above_threshold)/nrow(region_data))
}

pred_error_data <- sapply(regions, error_at_above_pred_error_cutoff, threshold=0.999)

pred_error_reduction <- 100 * (1 - pred_error_data["error",]/basic_rates)
pred_nseqs_reduction <- 100 * (1 - pred_error_data["nseqs",])
```

```{r oligos_coverage_analysis}
get_oligos_coverage <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good0 <- region_data$pbdiffs <= 0 & region_data$coverage >= 10
	region_data_good0 <- region_data[good0,]

	good1 <- region_data$pbdiffs <= 1 & region_data$coverage >= 10
	region_data_good1 <- region_data[good1,]

	c(	100*mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data),
		100*mean(region_data_good1[,"error"]), nrow(region_data_good1)/nrow(region_data))
}

oligos_coverage_error <- matrix(unlist(lapply(regions, get_oligos_coverage)), ncol=4, byrow=T)
rownames(oligos_coverage_error) <- regions
colnames(oligos_coverage_error) <- c("0.10.error", "0.10.frac", "1.10.error", "1.10.frac")

#     0.10.error 0.10.frac 1.10.error 1.10.frac
#V1V3  0.4515344 0.5898447  0.4632910 0.7968639
#V1V5  0.4187311 0.5557715  0.4270938 0.7384810
#V1V6  0.4399351 0.5410259  0.4460726 0.7167325
#V1V9  0.4317629 0.4444776  0.4420337 0.5752849
#V3V5  0.6421004 0.6854427  0.6555240 0.8102832
#V4    0.6825316 0.5512332  0.7023460 0.8245533

oligos_coverage_error_reduction <- 100 * (1 - oligos_coverage_error[,"1.10.error"]/basic_rates)
oligos_coverage_nseqs_reduction <- 100 * (1 - oligos_coverage_error[,"1.10.frac"])
```

```{r oligosQScoreAnalysis}
get_oligos_pred_error <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good0 <- region_data$pbdiffs <= 0 & region_data$pred_error >= 0.999
	region_data_good0 <- region_data[good0,]

	good1 <- region_data$pbdiffs <= 1 & region_data$pred_error >= 0.999
	region_data_good1 <- region_data[good1,]

	c(	100 * mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data),
		100 * mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data) )
}

oligos_pred_error <- matrix(unlist(lapply(regions, get_oligos_pred_error)), ncol=4, byrow=T)
rownames(oligos_pred_error) <- regions
colnames(oligos_pred_error) <- c("0.999.error", "0.999.frac", "1.999.error", "1.999.frac")

#     0.999.error 0.999.frac 1.999.error 1.999.frac
#V1V3   0.2806049  0.4749822   0.2806049  0.4749822
#V1V5   0.2177965  0.4266279   0.2177965  0.4266279
#V1V6   0.2168503  0.4296550   0.2168503  0.4296550
#V1V9   0.2177073  0.3438720   0.2177073  0.3438720
#V3V5   0.2229757  0.5131411   0.2229757  0.5131411
#V4     0.3485654  0.3961328   0.3485654  0.3961328

oligos_pred_error_reduction <- 100 * (1 - oligos_pred_error[,"1.999.error"]/basic_rates)
oligos_pred_nseqs_reduction <- 100 * (1 - oligos_pred_error[,"1.999.frac"])
```

```{r coverage_pred_error_analysis}
get_coverage_pred <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good <- region_data$coverage >= 10 & region_data$pred_error >= 0.999
	region_data_good <- region_data[good,]

	c(100 * mean(region_data_good[,"error"]), nrow(region_data_good)/nrow(region_data) )
}

coverage_pred_error <- matrix(unlist(lapply(regions, get_coverage_pred)), ncol=2, byrow=T)
rownames(coverage_pred_error) <- regions
colnames(coverage_pred_error) <- c("10.999.error", "10.999.frac")

#    10.999.error 10.999.frac
#V1V3   0.2843963    65.86436
#V1V5   0.2188792    56.55108
#V1V6   0.2178326    57.33756
#V1V9   0.2191284    42.66675
#V3V5   0.2236951    60.74254
#V4     0.3527900    61.79213

coverage_pred_error_reduction <- 100 * (1 - coverage_pred_error[,"10.999.error"]/basic_rates)
coverage_pred_nseqs_reduction <- 100 * (1 - coverage_pred_error[,"10.999.frac"])
```

```{r all_filters_analysis}
get_all_filters <- function(region){
	region_data <- mock_error_basic[mock_error_basic$region == region,]

	good0 <- region_data$pbdiffs == 0 & region_data$pred_error >= 0.999 & region_data$coverage >= 10
	region_data_good0 <- region_data[good0,]

	good1 <- region_data$pbdiffs <= 1 & region_data$pred_error >= 0.999 & region_data$coverage >= 10
	region_data_good1 <- region_data[good1,]

	c(	100 * mean(region_data_good0[,"error"]), nrow(region_data_good0)/nrow(region_data),
		100 * mean(region_data_good1[,"error"]), nrow(region_data_good1)/nrow(region_data) )
}

all_filters_error <- matrix(unlist(lapply(regions, get_all_filters)), ncol=4, byrow=T)
rownames(all_filters_error) <- regions
colnames(all_filters_error) <- c("0.10.999.error", "0.10.999.frac", "1.10.999.error", "1.10.999.frac")

#    0.10.60.error 0.10.60.frac 1.10.60.error 1.10.60.frac
#v4    0.001515500    0.5562125   0.001993325    0.7499762
#v13   0.009299146    0.4337597   0.009827404    0.6155779
#v35   0.008613200    0.4862353   0.008979207    0.6253128
#v15   0.003078476    0.3488609   0.003905562    0.5268616
#v16   0.007738039    0.2954687   0.008814923    0.4295469
#v19   0.004797594    0.2496316   0.005266467    0.3631005

#     0.10.999.error 0.10.999.frac 1.10.999.error 1.10.999.frac
#V1V3      0.2808218     0.4696170      0.2840387     0.6183602
#V1V5      0.2184805     0.4192252      0.2182989     0.5374066
#V1V6      0.2170989     0.4182253      0.2174090     0.5421862
#V1V9      0.2176989     0.3213802      0.2191334     0.4016057
#V3V5      0.2241187     0.5038275      0.2240779     0.5841414
#V4        0.3492870     0.3932393      0.3526755     0.5665274

all_filters_error_reduction <- 100 * (1 - all_filters_error[,"1.10.999.error"]/basic_rates)
all_filters_nseqs_reduction <- 100 * (1 - all_filters_error[,"1.10.999.frac"])
```

```{r summary}
base_error_reduction <- rbind(	pb_error_reduction_0,
								pb_error_reduction_1,
								pred_error_reduction,
								coverage_error_reduction,
								coverage_pred_error_reduction,
								oligos_coverage_error_reduction,
								oligos_pred_error_reduction,
								all_filters_error_reduction)

base_nseqs_reduction <- rbind(	pb_nseqs_reduction_0,
								pb_nseqs_reduction_1,
								pred_nseqs_reduction,
								coverage_nseqs_reduction,
								coverage_pred_nseqs_reduction,
								oligos_coverage_nseqs_reduction,
								oligos_pred_nseqs_reduction,
								all_filters_nseqs_reduction)
```

>>>>

@@

```{r random_stuff }

@@

#This is for the Figure 2C plot...
coverage_overall_agg <- aggregate(100 * mock_error_basic$error, by=list(mock_error_basic$coverage), mean)
coverage_overall <- coverage_overall_agg$x
names(coverage_overall) <- coverage_overall_agg$Group.1
coverage_overall <- coverage_overall[as.character(3:40)]

#This is for the Figure 2D plot...
pred_overall <- aggregate(100 * mock_error_basic$error, by=list(round(mock_error_basic$pred_error, 3)), mean)
